---
title: "Task2"
author: "Bao Luu"
date: "2025-06-29"
output: html_document
---

# 1. Install & Import libraries

```{r}
library(tidyverse)
library(dplyr)
library(psych)
```

# 2. Read & Understand dataset

```{r}
df2 <- read.csv("../data/diabetes.csv", 
               encoding = "latin1",
               fileEncoding = "latin1")
head(df2)
```

```{r}
sapply(df2, class)
```

```{r}
colSums(is.na(df2))
```

# 3. Task d

```{r}
desc_stats <- function(data_column) {
  if (is.numeric(data_column)) {
    mean_val <- mean(data_column, na.rm = TRUE)
    median_val <- median(data_column, na.rm = TRUE)
    sd_val <- sd(data_column, na.rm = TRUE)
    var_val <- var(data_column, na.rm = TRUE)
    return(c(mean = mean_val, median = median_val, standard_deviation = sd_val, variance = var_val))
  } else {
    return(c(mean = NA, median = NA, standard_deviation = NA, variance = NA))
  }

}
```

```{r}
for (col_name in names(df2)){
  cat(paste("Statistics of ", col_name, "\n"))
  stat_names <- desc_stats(df2[[col_name]])
  for (stat_name in names(stat_names)) {
    cat(stat_name,":", stat_names[[stat_name]]," ")
  }
  cat("\n")
}

```

```{r}
num_cols = c
for (col_name in names(df2)){
  plot <- ggplot(df2, aes_string(x = col_name)) +
    geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
    labs(
      title = paste("Distribution of", col_name, "before cleaning"),
      x = col_name,
      y = "Frequency (Count)"
    ) +
    theme_light()
  
  print(plot)
}
```

```{r}
draw_boxplot <- function(column, title) {
  boxplot(column, main = title, horizontal = TRUE)
}

for (col_name in names (df2)){
  if (is.numeric(df2[[col_name]])){
    title_text <- paste("Descriptive stat of", col_name, "before cleaning")
    draw_boxplot(df2[[col_name]], title = title_text)
    cat("\n")
  }
}
```

# 4. Task e

### Errors Identified in Each Column

### Glucose

-   **Problem**: 0 mg/dL is not physiologically possible

-   **Normal Range**: 70–100 mg/dL (fasting)

```{r}
df2 %>% 
  filter(Glucose == 0)
```

### Blood Pressure

-   **Problem**: 0 mm Hg diastolic pressure is unrealistic.

-   **Normal Range**: 60–80 mm Hg (diastolic).

```{r}
df2 %>% 
  filter(BloodPressure == 0)
```

### SkinThickness

-   **Problem**: 0 mm triceps skinfold is impossible.

-   **Normal Range**: 10–50 mm in adults.

```{r}
df2 %>% 
  filter(SkinThickness == 0)
```

### Insulin

-   **Problem**: 0 µU/mL is highly unlikely for serum insulin.

-   **Normal Range**: 2–25 µU/mL fasting.

```{r}
df2 %>% 
  filter(Insulin == 0)
```

### BMI

-   **Problem**: BMI = 0 means zero weight — impossible.

-   **Normal Range**: 18.5–24.9 (normal weight).

```{r}
df2 %>% 
  filter(BMI == 0)
```

```{r}
cols_to_clean <- c("Glucose", "BloodPressure", "SkinThickness", "Insulin", "BMI")
for (col in cols_to_clean) {
  df2[[col]][df2[[col]] == 0] <- NA
}

for (col in cols_to_clean) {
  df2[[col]][is.na(df2[[col]])] <- mean(df2[[col]], na.rm = TRUE)
}
```

```{r}
describeBy(df2)
```

```{r}
for (col_name in names(df2)){
  plot <- ggplot(df2, aes_string(x = col_name)) +
    geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
    labs(
      title = paste("Distribution of", col_name, "after cleaning"),
      x = col_name,
      y = "Frequency (Count)"
    ) +
    theme_light()
  
  print(plot)
}
```

```{r}
for (col_name in names (df2)){
  if (is.numeric(df2[[col_name]])){
    title_text <- paste("Descriptive stat of", col_name, "after cleaning")
    draw_boxplot(df2[[col_name]], title = title_text)
    cat("\n")
  }
}
```
